---
title: "Thesis synthesis code"
output:
  html_document:
    #code_folding: hide
    depth: 1
    number_sections: no
    theme: yeti
    toc: yes
    toc_float: yes
    code_download: true
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, cache = TRUE)
```

## Load packages

```{r}
library(tidyverse)
library(R.utils) # to unzip the gzb files

```

## Load the data

I use `Number of live births` and `Number of children fathered` form the UK biobank as measures of lifetime reproductive success (LRS) in females and males respectively. 

```{r}
female_LRS <- read_tsv("Data/Summary/2734.gwas.imputed_v3.female.tsv")

male_LRS <- read_tsv("Data/Summary/2405.gwas.imputed_v3.male.tsv")
```

## Clean the GWAS SNP datasets

```{r}
female_clean <-
  female_LRS %>% 
  filter(low_confidence_variant == FALSE & minor_AF > 0.001) %>%
  select(variant, beta) %>% 
  rename(beta_female = beta)

male_clean <-
  male_LRS %>% 
  filter(low_confidence_variant == FALSE & minor_AF > 0.001) %>% 
  select(variant, beta) %>% 
  rename(beta_male = beta) 


data <-
left_join(female_clean, male_clean, by = "variant") %>% 
  filter(!is.na(beta_female) & !is.na(beta_male)) %>%  # remove NAs
  arrange(beta_female) %>%
  mutate(bin = c(rep(1:floor(n()/1000), each = 1000), 
                 rep(floor(n()/1000) + 1, each = n() %% 1000))) %>%
  group_by(bin) %>%
  summarise(females = mean(beta_female), males = mean(beta_male))

```

## Build Figure 1.1

```{r}

boyle_plot <- 
  data %>%
  ggplot(aes(females, males)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_vline(xintercept = 0, linetype = 2) +
  geom_point(alpha = 0.8) +
  stat_smooth(method = "lm", formula = y ~ x + I(x^2), size = 0.6) +
  #coord_cartesian(xlim = c(0, 0.25), ylim = c(0, 0.25)) +
  xlab("Mean effect size on female LRS in a human population\n(number of live births)") +
  ylab("Mean effect size on male LRS in a human population\n(number of children fathered)") + 
  theme_bw() +
  theme(strip.background = element_blank(),
        strip.text = element_text(hjust=0)) + 
  theme(text = element_text(size = 14))

boyle_plot
```

**Figure 1**. fitness is omnigenic. Each point represents the mean effect size for a group of 1000 variants, ordered by association with female lifetime reproductive success
